Language and vision in conceptual processing: Multilevel analysis and statistical power

### 2022-06-29 manuscript/manuscript.tex
29062022

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Language and vision in conceptual processing: Multilevel analysis and statistical power},
  pdflang={en-EN},
  pdfkeywords={semantic, conceptual processing, vocabulary size, statistical power},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{semantic, conceptual processing, vocabulary size, statistical power}
\usepackage{csquotes}
\renewcommand\author[1]{}
\renewcommand\affiliation[1]{}
\authorsnames[1, 2, 2]{Pablo Bernabeu, Dermot Lynott, Louise Connell}
\authorsaffiliations{{Department of Psychology, Lancaster University, UK}, {Department of Psychology, Maynooth University, Ireland}}
\renewcommand{\topfraction}{.75}
\renewcommand{\bottomfraction}{.4}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.6}
\setlength{\@fptop}{0pt}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\usepackage{amsmath}
\usepackage{upgreek}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\renewcommand\appendix{}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\abstract{%
Research over the past two decades has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the data: individual differences, words and tasks. To this aim, we drew on three existing data sets that were composed of large samples, and implemented the paradigms of semantic priming, semantic decision and lexical decision. After extending these data sets with language-based and vision-based measures, we performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. Overall, language-based information was found to be more important than vision-based information. Furthermore, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. In addition, higher-vocabulary participants presented a greater effect of language-based similarity than lower-vocabulary participants. Conversely, lower-vocabulary participants presented a greater effect of vision-based information, which was less relevant to the task. This `relevance advantage' for higher-vocabulary participants also appeared in the semantic decision study, whose task required distinguishing between concrete and abstract words. Specifically, higher-vocabulary participants were more strongly influenced by word concreteness than lower-vocabulary participants. Furthermore, we compared the effects of different measures of visual information. Lastly, we estimated the sample size required to reliably examine each effect of interest. We found that 300 participants would suffice to examine language-based information contained in words, whereas more than 1,000 participants would be necessary to examine vision-based information as well as the interactions between both the former variables and vocabulary size, gender and presentation speed.
}

@@

Research over the past two decades has suggested that conceptual processing involves both `linguistic' and `embodied' systems of the brain. The language system is characterised by associations across networks of words (Landauer et al., 1998; Pylyshyn, 1973), whereas the embodiment system is characterised by associations within perceptual, motor, affective and social domains (Barsalou, 1999; Diveica et al., 2022). Both systems have strengths and limitations (Connell, 2019; Dove, 2020; Kumar, 2021; Louwerse et al., 2015; Mahon \& Hickok, 2016). As a result, conceptual processing flexibly draws on these systems following the demands of the context (Sato et al., 2008; Shebani \& Pulvermüller, 2018; Willems \& Casasanto, 2011).

The language system is involved in the representation of both abstract and concrete concepts (Ponari, Norbury, Rotaru, et al., 2018; Reggin et al., 2021). For instance, measures of distributional semantic distance---solely based on texts---have been shown to capture aspects of the real world with considerable accuracy (Contreras Kallens et al., 2022; De Deyne et al., 2013; Kumar et al., 2020; Louwerse \& Zwaan, 2009). Although the language system is likely more important for the processing of abstract---rather than concrete---concepts (Barca et al., 2020; Duñabeitia et al., 2009; Hultén et al., 2021), although the degree of this difference could be small (De Deyne et al., 2013). Language-based regularities may even allow distinguishing between concrete and abstract words, as word concreteness is consistently related to linguistic characteristics such as part of speech, morphological structure, countability and etymology (Lievers et al., 2021).

Regarding the time course, the language system is activated more quickly in semantically-deep tasks (Connell \& Lynott, 2013; Louwerse \& Connell, 2011), whereas both systems are similarly fast in semantically-shallow tasks (Petilli et al., 2021). As a result of these characteristics, the language system is equipped to support a major part of conceptual processing (Connell, 2019; Connell \& Lynott, 2014). However, the language system is hindered by the symbol grounding problem, which refers to the impossibility of obtaining meaning---or `grounding' lexical content in reality---on the sole basis of word-to-word associations (Günther et al., 2021, 2018; Harnad, 1990; Louwerse, 2011). Another limitation is caused by the finite vocabularies of languages, which sometimes fall short of the richness of sensorimotor experience (De Deyne et al., 2021; Günther et al., 2021; Majid \& Burenhult, 2014; Majid \& Levinson, 2011; Majid \& van Staden, 2015). In conclusion, word-to-word associations appear to be necessary but not sufficient for conceptual processing.

The embodiment system is best suited for the representation of concrete concepts, such as those involving perceptual features (Ostarek \& Huettig, 2017; Pecher et al., 1998; Yee et al., 2012) and actions (Riccardi et al., 2019). Some research has suggested that embodied simulation plays an integral role in conceptual processing (Pulvermüller, 2013), at least in the context of relatively-semantic tasks and stimuli (Connell \& Lynott, 2016; García et al., 2020; Lebois et al., 2015; Ostarek \& Huettig, 2017; Pecher et al., 1998; Shebani et al., 2021; Vitale et al., 2021; Vukovic et al., 2017; X. Wang et al., 2021; Yee et al., 2012). For instance, Riccardi et al. (2019) investigated the comprehension of action verbs and abstract verbs in left-hemisphere stroke patients, and found that the processing of action concepts relied more heavily on sensory-motor areas. Furthermore, studies have suggested that even abstract concepts may be partially grounded in modality-specific domains such as emotion, sociality and interoception (Conca et al., 2021; Connell et al., 2018; Diveica et al., 2022; Kousta et al., 2011; Vigliocco et al., 2014; X. Wang et al., 2021).

However, other research has curtailed the importance of this system (Bottini et al., 2021; Garcea et al., 2013; Günther et al., 2021, 2018; Mahon \& Caramazza, 2008; Montero-Melis et al., 2022; Vannuscorps et al., 2016). For instance, studies have found that lesions to motor brain areas did not systematically impair the processing of action concepts in language (Negri et al., 2007; Papeo et al., 2010).

\hypertarget{interplay-between-language-and-embodiment-systems}{%
\subsection{Interplay between language and embodiment systems}\label{interplay-between-language-and-embodiment-systems}}

The above characteristics make the language and the embodiment systems complementary to each other (Andrews et al., 2009; De Deyne et al., 2021; J. Wang et al., 2010). A comprehensive approach to both these systems requires a direct comparison of them. Studies tackling such a comparison have found that the systems are selectively engaged, following contextual demands (Connell \& Lynott, 2013; Louwerse \& Connell, 2011; Ostarek \& Huettig, 2017; Petilli et al., 2021). For instance, the role of perceptual simulation appears to be augmented in slower responses (Louwerse \& Connell, 2011), when words are presented more slowly (Lam et al., 2015), and in tasks prompting deeper semantic processing (Connell \& Lynott, 2013; Louwerse \& Connell, 2011; Ostarek \& Huettig, 2017; Petilli et al., 2021).

\hypertarget{the-present-study}{%
\section{The present study}\label{the-present-study}}

There is considerable evidence supporting the `hybrid' theory positing that conceptual processing depends on both language and embodiment systems (Barsalou et al., 2008; Connell \& Lynott, 2013; Louwerse, 2011). Nonetheless, there are also reasons to continue testing this theory. Firstly, the coexistence of several systems in a scientific theory must be thoroughly justified, due to the value of simplicity, as epitomised by Occam's razor (Gallese \& Lakoff, 2005; Tillman et al., 2015). This point is particularly pressing because the language system has consistently produced larger effect sizes than the embodiment system (Banks et al., 2021; Kiela \& Bottou, 2014; Lam et al., 2015; Louwerse et al., 2015; Pecher et al., 1998; Petilli et al., 2021). Secondly, mixed evidence has appeared regarding some aspects of the interplay. For instance, whereas some studies have suggested that the language system becomes relevant before the embodiment system (Lam et al., 2015; Louwerse \& Connell, 2011), a recent study found the opposite pattern (Petilli et al., 2021). Similarly, whereas some evidence has suggested that female participants might draw on the language system more prominently than males (Hutchinson \& Louwerse, 2013), other research has suggested that this difference could be negligible (Wallentin, 2020). Thirdly, the role of individual differences has not been examined in most of the studies addressing the language--embodiment interplay. Fourthly and last, the scarcity of statistical power that has ailed some studies in cognitive psychology and neuroscience (Marek et al., 2022; Vasishth \& Gelman, 2021) is likely to have affected some studies in the present topic area, too (see Lynott et al., 2014; Montero-Melis et al., 2022).

The above open questions were addressed in the present study. These questions were overarched by the main research question, which examined whether the interplay between linguistic and visual processing held in the expected direction and at multiple levels of the experimental structure. To study this question, we drew on the data sets from three previous studies whose samples of participants were larger than average. These studies implemented the paradigms of semantic priming (Hutchison et al., 2013), semantic decision (Pexman et al., 2017) and lexical decision (Balota et al., 2007). Importantly, we extended these data sets with

@@

Experimental data in psycholinguistics often consists of various levels, such as participants, words and tasks. A simultaneous examination of these levels should improve our understanding of the evidence collected (Ostarek \& Bottini, 2021). For instance, this cross-examination can help increase our certainty on the attribution of explanatory power to each level, as well as allowing a comparison of the explanatory power across levels. This multilevel approach is complementary to another approach in the field which seeks to test the causal contribution of different sources of information to conceptual processing---e.g., language (Ponari, Norbury, Rotaru, et al., 2018), perception (Stasenko et al., 2014) and action (Speed et al., 2017).

@@

\hypertarget{participant-level}{%
\subsection{Participant level}\label{participant-level}}

*** to be edited *** Individual differences in domains such as language, perception, mental imagery and physical experience (e.g., Davies et al., 2017; Dils \& Boroditsky, 2010; Fetterman et al., 2018; Holt \& Beilock, 2006; Mak \& Willems, 2019; Miceli et al., 2022; Pexman \& Yap, 2018; Vukovic \& Williams, 2015; Yap et al., 2012, 2017).\footnote{According to Lamiell (2019), `individual differences' is a misnomer in that the methods used to analyse individual differences (e.g, regression) are not participant-specific. We think that this observation is mitigated in the context of mixed-effects models (as used in our current study), which incorporate by-participant random intercepts and random slopes.} Furthermore, in some topic areas in which individual differences have not featured as prominently, the data have revealed a non-trivial role of individual differences (Kos et al., 2012; Montero-Melis, 2021).

@@

Vocabulary size refers to the number of words a person can recognise out of a discrete set. Higher-vocabulary participants are expected to respond faster overall (Pexman \& Yap, 2018). Some previous studies have found that the effect of vocabulary size was moderated by variables related to general intelligence, such as processing speed (Ratcliff et al., 2010; Yap et al., 2012). Owing to those findings, and the recommendations from other studies (James et al., 2018; Pexman \& Yap, 2018; Sassenhagen \& Alday, 2016), the present studies included covariates of vocabulary size, where available.These covariates were measures of intelligence that were not vocabulary-based, namely, attentional control in Study 1 and information uptake in Study 2 (such a variable was not available for Study 3).

@@

*** to be edited *** Semantic or lexical information in words (e.g., De Deyne et al., 2021; Lam et al., 2015; Lund et al., 1995; Lund \& Burgess, 1996; Lynott et al., 2020; Mandera et al., 2017; Pexman et al., 2017; Santos et al., 2011; Wingfield \& Connell, 2022)

\begin{itemize}
\item
  \emph{Variables in this study:} degree of association between words, degree of vision-based information in single words
\item
  Covariates: lexical variables and word concreteness.
\item
  Lexical variables: The lexical covariates were selected in each study out of the same 5 variables, which have been identified as important in previous research {[}Wingfield and Connell (2022); see \protect\hyperlink{general-statistical-analyses}{\underline{General statistical analysis}} for details{]}.
\item
  Word concreteness: Adjusting for word concreteness was deemed important due to the pervasive effect of this variable across lexical-semantic tasks (Brysbaert et al., 2014; Connell \& Lynott, 2012; Pexman \& Yap, 2018). Furthermore, it has been recently suggested that the effect of word concreteness does not stem from perceptual simulation, but instead from modality-independent properties of words (Bottini et al., 2021).
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  \emph{Variable in this study:} stimulus-onset asynchrony (SOA), present in Study 1 only.\footnote{The names of all variables were slightly adjusted for this text to facilitate their understanding---for instance, by replacing underscores with spaces (see code scripts at \url{https://osf.io/ueryq}). One specific case deserves further comment. We use the formula of the `stimulus-onset asynchrony' (SOA) in this paper, instead of the `interstimulus interval' (ISI), as the SOA has been more commonly used in previous papers (e.g., Hutchison et al., 2013; Pecher et al., 1998; Petilli et al., 2021; Yap et al., 2017). The difference between these formulas is that the ISI does not count the presentation of the prime word (example equivalences between ISI and SOA are available in Lam et al. (2015), and in Figure 1A of Di Lollo et al. (2004)). In the current study, the presentation of the prime word lasts 150 ms. Thus, the 200-ms SOA is equivalent to an ISI of 50 ms, and the 1,200-ms SOA corresponds to an ISI of 1,050 ms (Hutchison et al., 2013). The use of either formula in the analysis would not change our results, as we recoded the levels of the factor as -0.5 and +0.5, and then z-scored those, following the advice of Brauer and Curtin (2018). In our analyses (\url{https://osf.io/ueryq}), we used the ISI formula as it was the one present in the data set of Hutchison et al. (2013) (retrieved from \url{https://www.montana.edu/attmemlab/documents/all\%20ldt\%20subs_all\%20trials3.xlsx}}).
\end{itemize}

@@

The three levels are strongly intertwined, and studies have probed into more than one level at once: for instance, word level and individual level (Aujla, 2021; Lim et al., 2020; Pexman \& Yap, 2018; Yap et al., 2009), or word level and task level (Connell \& Lynott, 2013; Ostarek \& Huettig, 2019; Petilli et al., 2021). Thus, hypotheses are available regarding interactions across levels, as addressed below.

@@

\hypertarget{general-statistical-analyses}{%
\subsection{General method for statistical analyses}\label{general-statistical-analyses}}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
  man,floatsintext]{apa7}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Language and vision in conceptual processing: Multilevel analysis and statistical power},
  pdflang={en-EN},
  pdfkeywords={conceptual processing, semantic priming, semantic decision, lexical decision, language, vision, vocabulary size, statistical power},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{conceptual processing, semantic priming, semantic decision, lexical decision, language, vision, vocabulary size, statistical power}
\usepackage{csquotes}
\renewcommand\author[1]{}
\renewcommand\affiliation[1]{}
\authorsnames[1, 2, 2]{Pablo Bernabeu, Dermot Lynott, Louise Connell}
\authorsaffiliations{{Department of Psychology, Lancaster University, UK}, {Department of Psychology, Maynooth University, Ireland}}
\renewcommand{\topfraction}{.75}
\renewcommand{\bottomfraction}{.4}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.6}
\setlength{\@fptop}{0pt}
\setcounter{topnumber}{3}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\usepackage{amsmath}
\usepackage{upgreek}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\renewcommand\appendix{}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\abstract{%
Research over the past two decades has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the data: individuals, words and tasks. To this aim, we drew on three existing data sets that were composed of large samples, and implemented the paradigms of semantic priming, semantic decision and lexical decision. After extending these data sets with language-based and vision-based measures, we performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. Overall, language-based information was found to be more important than vision-based information. Furthermore, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. In addition, higher-vocabulary participants presented a greater effect of language-based similarity than lower-vocabulary participants. In contrast, this pattern did not hold for vision-based information, which was less relevant to the task. The `relevance advantage' in higher-vocabulary participants also appeared in the semantic decision and the lexical decision studies. Furthermore, we compare the effects of two visual information measures, and discuss the role of measurement instruments in this research topic. Lastly, we estimated the sample size required to reliably examine each effect of interest. We found that 300 participants suffice to examine language-based information contained in words, whereas more than 1,000 participants are necessary to examine vision-based information as well as the interactions between both the former variables and vocabulary size, gender and presentation speed.
}

@@

Over the past decades, research in psycholinguistics has suggested that conceptual processing depends on both language and embodiment systems (Barsalou et al., 2008; Connell \& Lynott, 2013; Louwerse, 2011). A comprehensive approach to both these systems requires a direct comparison of them. Studies tackling such a comparison have found that the systems are selectively engaged, following contextual demands (Connell \& Lynott, 2013; Louwerse \& Connell, 2011; Ostarek \& Huettig, 2017; Petilli et al., 2021). For instance, the role of perceptual simulation appears to be augmented in slower responses (Louwerse \& Connell, 2011), when words are presented more slowly (Lam et al., 2015), and in tasks prompting deeper semantic processing (Connell \& Lynott, 2013; Louwerse \& Connell, 2011; Ostarek \& Huettig, 2017; Petilli et al., 2021).

In spite of the amount of evidence favouring the interplay between language and embodiment, there are at least 2 reasons to continue testing the interplay theory. Firstly, the coexistence of several systems in a scientific theory must be thoroughly justified, due to the value of simplicity, as epitomised by Occam's razor (Gallese \& Lakoff, 2005; Tillman et al., 2015). This point is particularly pressing because the language system has consistently produced larger effect sizes than the embodiment system (Banks et al., 2021; Kiela \& Bottou, 2014; Lam et al., 2015; Louwerse et al., 2015; Pecher et al., 1998; Petilli et al., 2021). Consequently, in this research question, we examine whether the interplay between linguistic and visual processing holds in the expected directions and at multiple levels of the experimental structure.

Secondly, mixed evidence has appeared regarding some aspects of the interplay. For instance, whereas some studies have suggested that the language system is activated before the embodiment system (Lam et al., 2015; Louwerse \& Connell, 2011), a recent study found the opposite pattern in a lexical decision task (Petilli et al., 2021). Similarly, some evidence has suggested that high-vocabulary participants are more sensitive to linguistic features (Yap et al., 2017), whereas other evidence has suggested the opposite (Yap et al., 2009, 2012). Another case of mixed evidence regards gender: whereas some evidence has suggested that female participants draw on the language system more prominently than males (Hutchinson \& Louwerse, 2013), other research has suggested that this difference is negligible in the general population (Wallentin, 2020). Presumably, the scarcity of statistical power that has ailed some studies in cognitive psychology and neuroscience (Marek et al., 2022; Vasishth \& Gelman, 2021) may have affected some studies in the present topic area as well (see Lynott et al., 2014; Montero-Melis et al., 2022). Therefore, in this study, we re-examine longstanding questions using larger-than-average data sets, and calculate the sample size required to reliably detect a range of effects.

\hypertarget{language-and-vision}{%
\subsection{Language and vision}\label{language-and-vision}}

To retest the interplay theory, we revisit previous studies that contained larger-than-average samples of participants. Each of our three studies is organised around one \emph{hub} study. In Study 1, the hub is Hutchison et al. (2013), which implements a semantic priming paradigm. In Study 2, the hub is Pexman et al. (2017), which implements a lexical decision paradigm. In Study 3, the hub is Balota et al. (2007), which implements a lexical decision paradigm. Upon this basis, we added word-level variables from other studies to tap into language (Mandera et al., 2017; Wingfield \& Connell, 2022) and vision (Lynott et al., 2020; Petilli et al., 2021). Additionally, we included several covariates---or nuisance variables---to allow a rigorous analysis of the effects of interest (Sassenhagen \& Alday, 2016). These covariates comprised participant-specific variables (e.g., attentional control), lexical variables (e.g., word frequency) and word concreteness.

@@

Experimental data in psycholinguistics often consists of various levels, such as participants, words and tasks. A simultaneous examination of these levels should enhance our understanding of the evidence collected (Ostarek \& Bottini, 2021)---for instance, by shedding light on the distribution of explanatory power within and across levels. This multilevel approach is complementary to another approach in the field that seeks to test the causal contribution of different sources of information to conceptual processing---e.g., language (Ponari, Norbury, Rotaru, et al., 2018), perception (Stasenko et al., 2014) and action (Speed et al., 2017).

@@

\hypertarget{individual-level}{%
\subsection{Individual level}\label{individual-level}}

Variation The role of individual differences in domains such as language, perception, mental imagery and physical experience (e.g., Davies et al., 2017; Dils \& Boroditsky, 2010; Fetterman et al., 2018; Holt \& Beilock, 2006; Mak \& Willems, 2019; Miceli et al., 2022; Pexman \& Yap, 2018; Vukovic \& Williams, 2015; Yap et al., 2012, 2017).\footnote{According to Lamiell (2019), `individual differences' is a misnomer in that the methods used to analyse individual differences (e.g, regression) are not participant-specific. We think that this observation is mitigated in the context of mixed-effects models (as used in our current study), which incorporate by-participant random intercepts and random slopes.} Furthermore, in some topic areas in which individual differences have not featured as prominently, the data have revealed a non-trivial role of individual differences (Kos et al., 2012; Montero-Melis, 2021).

@@

Vocabulary size refers to the number of words a person can recognise out of a discrete set. Higher-vocabulary participants are expected to respond faster overall (Pexman \& Yap, 2018). Some previous studies have found that the effect of vocabulary size was moderated by variables related to general intelligence, such as processing speed (Ratcliff et al., 2010; Yap et al., 2012). Owing to those findings, and the recommendations from other studies (James et al., 2018; Pexman \& Yap, 2018; Sassenhagen \& Alday, 2016), the present studies included covariates of vocabulary size, where available. These covariates were measures of intelligence that were not vocabulary-based, namely, attentional control in Study 1 and information uptake in Study 2 (such a variable was not available for Study 3).

@@

*** to be edited *** Semantic or lexical information in words (e.g., De Deyne et al., 2021; Lam et al., 2015; Lund et al., 1995; Lund \& Burgess, 1996; Lynott et al., 2020; Mandera et al., 2017; Petilli et al., 2021; Pexman et al., 2017; Santos et al., 2011; Wingfield \& Connell, 2022)

\begin{itemize}
\item
  Variables in this study: language-based and vision-based information;
\item
  Covariates: lexical variables and word concreteness;

  \begin{itemize}
  \item
    Lexical variables: The lexical covariates were selected in each study out of the same 5 variables, which have been identified as important in previous research (Wingfield \& Connell, 2022; see \protect\hyperlink{general-statistical-analyses}{\underline{General statistical analysis}} for details).
  \item
    Word concreteness: Adjusting for word concreteness was deemed important due to the pervasive effect of this variable across lexical-semantic tasks (Brysbaert et al., 2014; Connell \& Lynott, 2012; Pexman \& Yap, 2018). Furthermore, it has been recently suggested that the effect of word concreteness does not stem from perceptual simulation, but instead from modality-independent properties of words (Bottini et al., 2021).
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  \emph{Variable in this study:} stimulus-onset asynchrony (SOA), present in Study 1 only.\footnote{The names of all variables were slightly adjusted for this text to facilitate their understanding---for instance, by replacing underscores with spaces (see code scripts at \url{https://osf.io/ueryq}). One specific case deserves further comment. We use the formula of the `stimulus-onset asynchrony' (SOA) in this paper, instead of the `interstimulus interval' (ISI), as the SOA has been more commonly used in previous papers (e.g., Hutchison et al., 2013; Pecher et al., 1998; Petilli et al., 2021; Yap et al., 2017). The difference between these formulas is that the ISI does not count the presentation of the prime word Di Lollo et al. (2004). In the current study, the presentation of the prime word lasts 150 ms. Thus, the 200-ms SOA is equivalent to an ISI of 50 ms, and the 1,200-ms SOA corresponds to an ISI of 1,050 ms (Hutchison et al., 2013). The use of either formula in the analysis would not change our results, as we recoded the levels of the factor as -0.5 and +0.5, and then \(z\)-scored those, following the advice of Brauer and Curtin (2018). In our analyses (\url{https://osf.io/ueryq}), we used the ISI formula as it was the one present in the data set of Hutchison et al. (2013) (retrieved from \url{https://www.montana.edu/attmemlab/documents/all\%20ldt\%20subs_all\%20trials3.xlsx}}).
\end{itemize}

@@

The three levels are strongly intertwined, and studies have probed into more than one level at once: for instance, word level and individual level (Aujla, 2021; Lim et al., 2020; Pexman \& Yap, 2018; Yap et al., 2009), or word level and task level (Al-Azary et al., 2022; Connell \& Lynott, 2013; Ostarek \& Huettig, 2019; Petilli et al., 2021). Thus, hypotheses are available regarding interactions across levels, as addressed below.

@@

\hypertarget{general-statistical-analyses}{%
\subsection{General method for statistical analysis}\label{general-statistical-analyses}}
